Smart Spreading Factor Assignment for LoRaWANs

### 2019-04-28 paper.tex
decrease figure sizes
Add Smetech ADR as realted work

LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and air time of transmission \cite{7803607}. Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other. Thus, SF assignment is crucial for overall network performance. In this work, a LoRa discrete event simulator is developed to study the performance of LoRa SF assignment strategies. Also, a machine learning based SF assignment approach is proposed. Support Vector Machine (SVM) and Decision Tree Classifier (DTC) techniques are employed and the introduced schemes are called as smart SF schemes. The performance of the smart schemes is compared with the performance of the lowest SF assignment scheme. It is shown that the proposed smart SF assignment schemes give promising results.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision}
\caption{Collision between nodes close to the gateway.}
\label{fig:collision}
\end{figure}

@@

In \cite{8430542}, the authors investigated the impact of the interference caused by simultaneous LoRa transmissions with the same and different SFs. They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the gateway for modeling interference in multiple gateways scenarios. Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.

In \cite{8115779}, the authors introduced two new algorithms for SF assignment in LoRaWANs.
First algorithm assigns SFs based on the total number of connected devices, second algorithm assigns SFs by balancing air time of the packets in each SF. Based on the simulation results, it is shown that the proposed algorithms give promising results.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision_solution_single_gw}
\caption{Collision avoidance by using higher SF for nodes close to the gateway.}
\label{fig:collision_solution_single_gw}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision_solution_multi_gw}
\caption{Collision avoidance for intersecting gateways.}
\label{fig:collision_solution_multi_gw}
\end{figure}

@@

As for future work, transmit power optimization can be included to the proposed smart schemes. In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to gateways can decrease transmit power to save energy. This will make transmissions more vulnerable to interference thus requires extra care. Moreover, other machine learning methods can be investigated for SF assignment enhancement. Reinforcement learning could be a good candidate. Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve prediction performance.

>>>>

LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and air time of transmission \cite{7803607}. Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other. Thus, SF assignment is crucial for overall network performance. In this work, a LoRa discrete event simulator is developed to study the performance of LoRa SF assignment strategies. Also, a machine learning based SF assignment approach is proposed. Support Vector Machine (SVM) and Decision Tree Classifier (DTC) techniques are employed and the introduced schemes are called as smart SF schemes. The performance of the smart schemes are compared with the performance of the lowest SF assignment scheme. It is shown that the proposed smart SF assignment schemes give promising results.

@@

\begin{figure}
\centering
\includegraphics[width=0.85\linewidth]{collision.png}
\caption{Collision between nodes close to the gateway.}
\label{fig:collision}
\end{figure}

@@

\par In \cite{8267219} and \cite{8430542}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions. The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal. Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.

The adaptive data rate (ADR) algorithm recommended by Semtech Corporation utilizes signal to noise ratio (SNR) of the last 20 transmissions to minimize the transmit power and SF while ensuring successful transmissions \cite{lorawan_adr}. The recommended algorithm uses a predefined SF and corresponding required SNR table. The algorithm considers maximum SNR of the last 20 transmissions. The algorithm increases transmit power or decrease data rate in case of low SNR and the opposite in case of high SNR.

@@

\begin{figure}
\centering
\includegraphics[width=0.85\linewidth]{collision_solution_single_gw.png}
\caption{Collision avoidance by using higher SF for nodes close to the gateway.}
\label{fig:collision_solution_single_gw}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{collision_solution_multi_gw.png}
\caption{Collision avoidance for intersecting gateways.}
\label{fig:collision_solution_multi_gw}
\end{figure}

@@

As for future work, transmit power optimization can be included to the proposed smart scheme. In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to gateways can decrease transmit power to save energy. This will make transmissions more vulnerable to interference thus requires extra care. Moreover, other machine learning methods can be investigated for SF assignment enhancement. Reinforcement learning could be a good candidate. Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme in order to improve prediction performance.
